---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

An introduction of the data and a description of the trends/books/items you are choosing to analyze (and why!)

The Seattle Public Library has available checkout data from 2005 to present. This data set provides a variety of information about checkouts for physical and electronic items. In this report, I am using a subset of this data, which covers all items checked out in 2022-2023 so far. I am focusing on exploring trends pertaining to checkouts for popular materials, publishers, and usage classes over the course of this specific time period. I chose these trends because they provide valuable insights about the preferences of Seattle Public Library patrons, as well as the materials and publishers that are most often checked out at SPL. Analyzing these trends can also provide insights about large-scale societal trends, such as shifts in the preferred medium of reading over the span of a year. This information can be useful not only for the SPL system, but also for researchers and policymakers analyzing literacy-related trends.

### Summary Information

From 2022-2023 so far, the Seattle Public Library had a total of 9,196,991 checkouts. Of these checkouts, 55.9% were digital materials and 44.1% were physical materials. These percentages can help the library better understand the preferences of their patrons and can improve future development decisions. The month with the highest number of checkouts was January 2023, with 269,816 checkouts, while the lowest number of checkouts was in December 2022, with 212,354 books checked out. This information can help the library identify popular usage times and plan staffing and resource allocation accordingly.


### The Dataset

- Who collected/published the data?  
- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
- How was the data collected or generated?  
- Why was the data collected?  
- What, if any, ethical questions do you need to consider when working with this data?  
- What are possible limitations or problems with this data?   (at least 200 words)


### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("A3Chart1.R")}
```

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("A3Chart2.R")}
```

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("A3Chart3.R")}
```


